Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-11-09 submit/apj.tex
simulate covariance matrices with stan

This article presents the first application of a computationally intensive analysis, so although  high spectral resolution
is available we here compress the information into low-resolution broadband photometry. 
Synthetic supernova-frame photometry is generated for a top-hat filter system
comprised of five 
bands with the following wavelength ranges: ${\hat{U}}$ $[3300.00 - 3978.02]$\AA;
${\hat{B}}$ $[3978.02-4795.35]$\AA;
$\hat{V}$ $[4795.35-5780.60]$\AA;
$\hat{R}$ $[5780.60-6968.29]$\AA;
$\hat{I}$ $[6968.29-8400.00]$\AA.
(The diacritic hat serves as a reminder that these are not standard Johnson-Cousins filters.)
For each supernova, the magnitudes within 5-days of peak brightness are used to fit single-band magnitudes
at $B$-band peak brightness.
The equivalent widths of the Si~II~$\lambda 4141$ and Ca~II H\&K features are computed as
in \citet{2008A&A...477..717B} and the 
wavelength of the Si~II~$\lambda 6355$ feature
as in \citet{chotard:thesis, 2017Chotard}.
Equivalent widths and the
Si~II~$\lambda 6355$ wavelength are taken from spectra  within $\pm 2.5$ days from $B$-band maximum;
the average is used  in cases where there are multiple spectral measurements within that time window.
\color{red}
All the above spectral features have units of \AA ngstroms.
\color{black}

@@

fluxes to luminosites.

@@

\color{red}(with units of mag) and slopes (with units mag~\AA$^{-1}$)
\color{black} of the linear relationships that

@@

The global parameter vectors $\vec{\gamma}^0$, $\vec{\gamma}^1$  are the slopes (\color{red}
with units of mag) \color{black} that relate 

@@

Another form of degeneracy is illustrated by the  transformations $\Delta \rightarrow \Delta  + \epsilon_\Delta$,
 $\vec{\gamma}^0 \rightarrow \vec{\gamma}^0  + \epsilon_{\gamma^0}$, $\vec{\gamma}^1 \rightarrow \vec{\gamma}^1 + \epsilon_{\gamma^1}$
 under the condition
$$
\epsilon_\Delta  +  \epsilon_{\gamma^0} k_0+  \epsilon_{\gamma^1} k_1=0,
$$
i.e.\ shifts in the global coefficients $\vec{\gamma}$ can be compensated by shifts in 
\color{red}
$\Delta$
\color{black}
for each supernova while maintaining 
$\langle \Delta \rangle=0$.
Another degeneracy is illustrated by $k_0 \rightarrow k_0 + (EW_{Ca}-\langle EW_{Ca}\rangle)\epsilon$,
$\vec{\alpha} \rightarrow \vec{\alpha} - \vec{\gamma} \epsilon$, $\vec{c} \rightarrow \vec{c} + \langle EW_{Ca}\rangle \epsilon$, i.e. the
linear external terms can leak into the linear spectral feature corrections.
These degeneracies are not broken within the model, but rather are constrained by the priors in which the
per-supernova parameters
$\Delta$, $EW$, $\lambda$, $k_0$, and $k_1$ are uncorrelated.

The parameterized intrinsic dispersion, $C_c$,  introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\vec{\gamma}^0 k_0$, and $\vec{\gamma}^1 k_1$ can  be attributed to intrinsic dispersion
\color{blue}
and vice versa.
In the case where there is independent intrinsic dispersion in all 5 bands, i.e.\  the input $C_c$ has rank-5, there exists degenerate solutions with a $C_c$ of rank 2.  Those identical predictions generally do not yield the same likelihood, as the determinant of $C_c$ varies over those solutions.  In a maximum-likelihood analysis the rank-2 is preferred over the rank-5 solution, resulting in some intrinsic dispersion getting attributed to $\Delta$, $k_0$, and $k_1$.
In our Bayesian analysis the priors, though not totally disallowing rank-2 $C_c$, favor rank-5  and prevent convergence to this solution.    
Leakage from per-supernova intrinsic color parameters to intrinsic dispersion could bias the intrinsic color vectors of the SN Ia family,  or conversely, intrinsic dispersion could be spuriously attributed to an intrinsic color.  We examine this effect in the discussion of the results in \S\ref{model2:sec}.
\color{black}
%
%There are several features of the model
%that drive the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$
%\color{red}
%and hence favors solutions of $\gamma$'s that minimize the residual dispersion described by $C_c$;
%\color{black}
%The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
%be non-Gaussian, and so are not well described by the Normal covariance we impose in $C_c$; the grey magnitude offsets, $\Delta$, would appear as a constant
%in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

In our Bayesian analysis the priors must be described.  A flat prior is used for all parameters except
for the covariance matrix $C_c$.
\color{red}
This matrix is parameterized in two parts: by the Choleksy factor of its
\color{black}
correlation matrix with a $\nu=4$  LKJ prior\footnote{
Visualization of the LKJ correlation distribution can be found in \url{http://www.psychstatistics.com/2014/12/27/d-lkj-priors/}.}
\citep{Lewandowski20091989}; its standard
deviations $\sigma_i = \sqrt{C_{c,ii}}$ with a  Cauchy distribution prior with location
 $0.1$ and scale $0.1$ mag restricted to positive values.
This covariance matrix prior is recommended by STAN \citep{stan}, the Monte Carlo program we use to evaluate the model.
 We find that a simpler intrinsic dispersion model with no correlations between bands
 (i.e.\ a  diagonal covariance matrix) produces \color{red} negligible \color{black} change in the posteriors of
 the other parameters.

@@

Tests have revealed no evidence for the non-convergence of the MCMC run.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
(in this paragraph not to be confused with the synthetic $R$ magnitude)
\citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that all are densely sampled.
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig} -- \ref{global5:fig}.  In these tests there is no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
For four selected colors, a plot of the
\color{red}
model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
\color{black}
 versus observed colors and the histogram of their differences
are shown in Figure~\ref{residual:fig}. 
There is no evidence of catastrophically poor model predictions nor of extreme
outlying data.
We have rerun the analysis with a variety of initial conditions, including one with $\vec{\gamma}$ equal to zero, except for a small positive 
\color{red}
$\gamma^0_{\hat{U}}$ and small negative $\gamma^1_{\hat{U}}$;
\color{black}
the 68\% credible intervals
remain consistent.  Note that the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $k$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{residual.pdf} 
            \caption{
            \color{red}
            Model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
            versus observed colors, for four selected colors. 
            \color{black}
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

@@

The fit $\Delta$'s are highly correlated between supernovae.  To aid 
\color{red} visual interpretation, we remove much of this
correlation by
selecting an arbitrary supernova and subtracting out its $\Delta|_0$ from those of all other objects, at the level of each MCMC link. \color{black}
The histogram of the \color{red} per-supernova \color{black} medians of these relative  grey offsets $\Delta-\Delta|_0$,
is shown in Figure~\ref{hist:fig}.  The distribution
has a standard deviation of
%-----
$0.08$
%-----
mag, and a tail in the positive (fainter) direction.
The supernovae with the highest values of $\Delta$ have, with increasing $\Delta$, heliocentric
 redshifts of 0.0228, 0.0086, 0.0565, and 0.0015.  Peculiar velocities  can cause a significant magnitude offset for three.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{deltaDelta_hist.pdf} 
   \caption{
   Normalized \color{red} stack of the posteriors of all supernovae \color{black} and histogram of the \color{red} per-supernova \color{black} medians of the grey offset $\Delta$. 
   \color{red}  To help null correlated errors, we select
    an arbitrary supernova and subtract out its $\Delta|_0$ from those of all other objects at the level of each MCMC link.
   \color{black}
   The supernovae with the four highest values of $\Delta$ have, with increasing $\Delta$, heliocentric
   redshifts of 0.0228, 0.0086, 0.0565, and 0.0015.
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude
and color dispersions are captured in $C_c$
at the level of 4--10$\sigma$.
The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
\color{green}
The posterior of $\sigma_{\hat{R}}$ seen in Figure~\ref{global4:fig} does not exclude values that approach zero, highlighting that non-rank 5
$C_c$'s are accommodated by our prior.
\color{black}
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in ${\hat{V}}$ and colors with respect to ${\hat{V}}$ are
\begin{multline}
\sigma({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\left(
0.021^{+0.004}_{-0.005} ,\, 
0.067^{+0.010}_{-0.010},\, 
0.034^{+0.005}_{-0.005},\, 
0.023^{+0.002}_{-0.002},\, 
0.053^{+0.004}_{-0.004}
\right)
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\begin{pmatrix}
\begin{array}{rrrrr}
\hat{V} & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
\ldots & {\hat{U}}-{\hat{V}} & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
\ldots  & \ldots & {\hat{B}}-{\hat{V}} & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
\ldots  & \ldots  &\ldots  & {\hat{V}}-{\hat{R}} & 0.65^{+0.06}_{-0.07} \\
\ldots  & \ldots  &\ldots  &\ldots & {\hat{V}}-{\hat{I}} \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
To reduce clutter only the upper triangular part of the otherwise symmetric matrix is shown.
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
\citetalias{2017ApJ...842...93M}  find an intrinsic dispersion of the SALT2 $c$ parameter (associated with ${\hat{B}}-{\hat{V}}$) of $\approx 0.07$ mag,
though variation of dust properties accounts for some of that dispersion.
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eq.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and 
artificially
smaller statistical uncertainties.

@@

The above result is visualized in Figure~\ref{plane:fig}, which  shows  in the
\color{red} ${\hat{U}}{\hat{V}}\hat{I}$-subspace \color{black}
two perspectives (left and right panels)
of the unit vectors corresponding to $\vec{\gamma}^0$ and  $\vec{\gamma}^1$ of our model (solid lines),
and $\vec{a}$, $\vec{b}$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  All  vectors are set to intersect the origin.
The two perspectives show that while the four vectors point in different directions for each band combination,
they are almost coplanar in ${\hat{U}}{\hat{V}}{\hat{I}}$ (they are slightly less coplanar in ${\hat{B}}{\hat{V}}{\hat{R}}$).  The $\vec{a}$ and $\vec{b}$ vectors and the $\vec{\gamma}^0$ and $\vec{\gamma}^1$
vectors span near-parallel planes in the 5-dimensional color space, and hence the latter can be almost entirely expressed in terms of
the former.

@@

Given that
$\vec{\gamma_0}$  has nearly the same effect on color as dust extinction with constant $R^F$,
the distribution of
our parameter $\gamma^0_{\hat{V}} k_0$ can be compared with the expected distribution
of $A^F_V$.
Figure~\ref{k0_med:fig} shows the histogram of
\color{red}
per-supernova
\color{black}
median values
from the Monte Carlo chains
and the  \color{red} stack of the posteriors of all supernovae \color{black}  for $\gamma^0_{\hat{V}} k_0$ 
relative to that of an arbitrary supernova  $\gamma^0_{\hat{V}} k_0|_0$.
(Subtracting the value of the one supernova suppresses correlated errors in $k_0$ common to all supernovae.)
The distributions are non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with
simulations based on expected dust distributions
within galaxies and the distribution of galaxy orientations with respect to the observer,
which determine column densities along the lines of sight toward supernovae that are consistent with observations
\citep{1998ApJ...502..177H, 2007ApJ...659..122J}.  
We emphasize that unlike other analyses,
this distribution is given by the data, and so is not dependent on any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{deltagamma1_med.pdf}
      \caption{
      Normalized  \color{red} stack of the posteriors of all supernovae \color{black}  and histogram 
      of \color{red} per-supernova \color{black}
median values of: (left) $\gamma^0_{\hat{V}} k_0-\gamma^0_{\hat{V}} k_0|_0$, which is associated with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model;
(right) $\gamma^1_{\hat{V}} k_1-\gamma^1_{\hat{V}} k_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\langle R^F_{\mathit{eff}}\rangle$ value.
   \label{k0_med:fig}}
\end{figure}

We now consider the direction of $\vec{\gamma}^1$. 
Different supernovae are expected to encounter dust with different $R^F$;
having  associated $\vec{\gamma}^0$ with
the extinction from a single effective
$\langle R^F_{\mathit{eff}}\rangle$, $\vec{\gamma}^1$ is associated with perturbations
away from that effective value.
Let us introduce a new parameter for each supernova, $\delta R$, such that 
$A^F =\langle R^F_{\mathit{eff}}\rangle(1-  \delta R) E^F(B-V)$.
Variations in the value of  $\delta R$, with the prior of the non-correlation between 
the coefficients in Eq.~\ref{newdust:eqn},
lead 
to 
\begin{equation}
\kappa_2 =\langle R^F_{\mathit{eff}}\rangle \left(1- \frac{\mbox{Var}(\delta R \times E^F(B-V))}{\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))}\right).
\end{equation}
From the matrix elements of Eq.~\ref{trans_I:eqn},  $\kappa_2=-5.27/0.72=-7.31$, so that $\mbox{Var}(\delta R \times E^F(B-V)) = 4.00\times\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))$.  This result indicates that there is a small but significant covariance between  extinction contributions due to
fractional deviations from an effective $R^F$ and
the color-excess parameter $E^F(B-V)$.
Figure~\ref{k0_med:fig} shows the histogram of \color{red} per-supernova \color{black}
median values
from Monte Carlo chains
and  \color{red} stack of the posteriors of all supernovae \color{black}  of $\gamma^1_{\hat{V}} k_1$
relative to that of an arbitrary supernova  $\gamma^1_{\hat{V}} k_1|_0$.

@@

@@

The distributions of both sources of color excess $E({\hat{B}}-{\hat{V}})$ and
 $E_\delta({\hat{B}}-{\hat{V}}) \equiv   (\delta_{\hat{B}}-\delta_{\hat{V}})D$ are shown in Figure~\ref{ebv:fig} as the  \color{red} stack of the posteriors of all supernovae 
 and the histogram of per-supernova median values. \color{black} 
 (These values are relative
to
 those of an arbitrary supernova in order to null out correlated errors.)
The range of ${\hat{B}}-{\hat{V}}$ colors is $> 5 \times$ larger for the external contribution than for the
intrinsic
component.  
The standard deviations of  the median $E_\gamma({\hat{B}}-{\hat{V}})$ and $E_\delta({\hat{B}}-{\hat{V}})$ are
%-----
0.075
and 0.014
%-----
mag respectively.
It should be noted that
the $E_\delta({\hat{B}}-{\hat{V}})$ standard deviation is comparable to its measurement uncertainty per-supernova.
Thus, since
the color excesses of individual supernovae have low signal-to-noise,
the significance of $\vec{\delta}$
is drawn from the ensemble of objects.

>>>>

This article presents the first application of a computationally intensive analysis, so although  high spectral resolution
is available we here compress the information into low-resolution broadband photometry. 
Synthetic supernova-frame photometry is generated for a top-hat filter system
comprised of five 
bands with the following wavelength ranges: ${\hat{U}}$ $[3300.00 - 3978.02]$\AA;
${\hat{B}}$ $[3978.02-4795.35]$\AA;
$\hat{V}$ $[4795.35-5780.60]$\AA;
$\hat{R}$ $[5780.60-6968.29]$\AA;
$\hat{I}$ $[6968.29-8400.00]$\AA.
(The diacritic hat serves as a reminder that these are not standard Johnson-Cousins filters.)
For each supernova, the magnitudes within 5-days of peak brightness are used to fit single-band magnitudes
at $B$-band peak brightness.
The equivalent widths of the Si~II~$\lambda 4141$ and Ca~II H\&K features are computed as
in \citet{2008A&A...477..717B} and the 
wavelength of the Si~II~$\lambda 6355$ feature
as in \citet{chotard:thesis, 2017Chotard}.
Equivalent widths and the
Si~II~$\lambda 6355$ wavelength are taken from spectra  within $\pm 2.5$ days from $B$-band maximum;
the average is used  in cases where there are multiple spectral measurements within that time window.
All the above spectral features have units of \AA ngstroms.

@@

fluxes to luminosities.

@@

with units of mag) and slopes (with units mag~\AA$^{-1}$)
of the linear relationships that

@@

The global parameter vectors $\vec{\gamma}^0$, $\vec{\gamma}^1$  are the slopes (with units of mag) that relate 

@@

Another form of degeneracy is illustrated by the  transformations $\Delta \rightarrow \Delta  + \epsilon_\Delta$,
 $\vec{\gamma}^0 \rightarrow \vec{\gamma}^0  + \epsilon_{\gamma^0}$, $\vec{\gamma}^1 \rightarrow \vec{\gamma}^1 + \epsilon_{\gamma^1}$
 under the condition
$$
\epsilon_\Delta  +  \epsilon_{\gamma^0} k_0+  \epsilon_{\gamma^1} k_1=0,
$$
i.e.\ shifts in the global coefficients $\vec{\gamma}$ can be compensated by shifts in 
$\Delta$
for each supernova while maintaining 
$\langle \Delta \rangle=0$.
Another degeneracy is illustrated by $k_0 \rightarrow k_0 + (EW_{Ca}-\langle EW_{Ca}\rangle)\epsilon$,
$\vec{\alpha} \rightarrow \vec{\alpha} - \vec{\gamma} \epsilon$, $\vec{c} \rightarrow \vec{c} + \langle EW_{Ca}\rangle \epsilon$, i.e. the
linear external terms can leak into the linear spectral feature corrections.
These degeneracies are not broken within the model, but rather are constrained by the priors in which the
per-supernova parameters
$\Delta$, $EW$, $\lambda$, $k_0$, and $k_1$ are uncorrelated.

The parameterized intrinsic dispersion, $C_c$,  introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\vec{\gamma}^0 k_0$, and $\vec{\gamma}^1 k_1$ can  be attributed to intrinsic dispersion
and vice versa.
In the case where there is independent intrinsic dispersion in all 5 bands, i.e.\  the input $C_c$ has rank-5, there exists degenerate solutions with a $C_c$ of rank 2.  Those identical predictions generally do not yield the same likelihood, as the determinant of $C_c$ varies over those solutions.  In a maximum-likelihood analysis the rank-2 is preferred over the rank-5 solution, resulting in some intrinsic dispersion getting attributed to $\Delta$, $k_0$, and $k_1$.
In our Bayesian analysis the priors, though not totally disallowing rank-2 $C_c$, favor rank-5  and prevent convergence to this solution.    
Leakage from per-supernova intrinsic color parameters to intrinsic dispersion could bias the intrinsic color vectors of the SN Ia family,  or conversely, intrinsic dispersion could be spuriously attributed to an intrinsic color.  We examine this effect in the discussion of the results in \S\ref{degenerate:sec}.

%
%There are several features of the model
%that drive the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$
%\color{red}
%and hence favors solutions of $\gamma$'s that minimize the residual dispersion described by $C_c$;
%\color{black}
%The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
%be non-Gaussian, and so are not well described by the Normal covariance we impose in $C_c$; the grey magnitude offsets, $\Delta$, would appear as a constant
%in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

In our Bayesian analysis the priors must be described.  A flat prior is used for all parameters except
for the covariance matrix $C_c$.
This matrix is parameterized in two parts: by the Choleksy factor of its
correlation matrix with a $\eta=4$  LKJ prior
\citep{Lewandowski20091989}; its standard
deviations $\sigma_i = \sqrt{C_{c,ii}}$ with a  Cauchy distribution prior with location
 $0.1$ and scale $0.1$ mag restricted to positive values.
This covariance matrix prior is recommended by STAN \citep{stan}, the Monte Carlo program we use to evaluate the model.
\color{blue}
Further properties of the this prior are presented in \S\ref{priorappendix:sec}.
\color{black}

@@

Tests have revealed no evidence for the non-convergence of the MCMC run.
STAN provides output statistics to assess
the convergence of the output Markov chains.
The 
potential scale reduction statistic, $\hat{R}$
(in this paragraph not to be confused with the synthetic $R$ magnitude)
\citep{Gelman92}, measures the convergence of the target distribution
in iterative simulations 
by using multiple independent sequences to estimate how much that distribution would sharpen if the simulations were run longer.
$N_{\mathit{eff}}$ is an estimate of the number of independent draws. The output for our data and model gives $\hat{R} \sim 1.0$ for all parameters, meaning there is no evidence for non-convergence.  The
output also gives  $N_{\mathit{eff}} \gg 100$ for all parameters, indicating that all are densely sampled.
Empirically, the confidence regions are localized and unimodal as is seen in  Figures~\ref{global1:fig} -- \ref{global5:fig}.  In these tests there is no evidence that
the Monte Carlo chains have not converged to the stationary posterior distribution.
For four selected colors, a plot of the
model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
 versus observed colors and the histogram of their differences
are shown in Figure~\ref{residual:fig}. 
There is no evidence of catastrophically poor model predictions nor of extreme
outlying data.
We have rerun the analysis with a variety of initial conditions, including one with $\vec{\gamma}$ equal to zero, except for a small positive 
$\gamma^0_{\hat{U}}$ and small negative $\gamma^1_{\hat{U}}$;
the 68\% credible intervals
remain consistent.  Note that the MCMC chains do not migrate to the parity-degenerate solution of $\vec{\gamma}$ and $k$.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{residual.pdf} 
            \caption{
            Model-expectation colors (including  supernova-specific spectral features and $\gamma k$ terms)
            versus observed colors, for four selected colors. 
            Plotted on the right for each color
is the histogram of the difference between the expected and
observed
 colors. 
            \label{residual:fig}}
\end{figure}

@@

The fit $\Delta$'s are highly correlated between supernovae.  To aid 
visual interpretation, we remove much of this
correlation by
selecting an arbitrary supernova and subtracting out its $\Delta|_0$ from those of all other objects, at the level of each MCMC link.
The histogram of the per-supernova  medians of these relative  grey offsets $\Delta-\Delta|_0$,
is shown in Figure~\ref{hist:fig}.  The distribution
has a standard deviation of
%-----
$0.08$
%-----
mag, and a tail in the positive (fainter) direction.
The supernovae with the highest values of $\Delta$ have, with increasing $\Delta$, heliocentric
 redshifts of 0.0228, 0.0086, 0.0565, and 0.0015.  Peculiar velocities  can cause a significant magnitude offset for three.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{deltaDelta_hist.pdf} 
   \caption{
   Normalized stack of the posteriors of all supernovae and histogram of the per-supernova medians of the grey offset $\Delta$. 
   To help null correlated errors, we select
    an arbitrary supernova and subtract out its $\Delta|_0$ from those of all other objects at the level of each MCMC link.
   The supernovae with the four highest values of $\Delta$ have, with increasing $\Delta$, heliocentric
   redshifts of 0.0228, 0.0086, 0.0565, and 0.0015.
   \label{hist:fig}}
\end{figure}

Non-trivial residual magnitude
and color dispersions are captured in $C_c$
at the level of 4--10$\sigma$.
The square
root of the diagonal elements are denoted as $\sigma$, which
range from
$\sim 0.01$ to 0.06 mag, significantly smaller
than the dispersion in $\Delta$.
Of particular interest within $C_c$ is intrinsic color dispersion.  The intrinsic dispersions in ${\hat{V}}$ and colors with respect to ${\hat{V}}$ are
\begin{multline}
\sigma({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\left(
0.021^{+0.004}_{-0.005} ,\, 
0.067^{+0.010}_{-0.010},\, 
0.034^{+0.005}_{-0.005},\, 
0.023^{+0.002}_{-0.002},\, 
0.053^{+0.004}_{-0.004}
\right)
 \text{mag}.
 \label{sig_intrinsic:eqn}
 \end{multline}
The correlations between these parameters are
\begin{multline}
Cor({\hat{V}},{\hat{U}}-{\hat{V}},{\hat{B}}-{\hat{V}},{\hat{V}}-{\hat{R}},{\hat{V}}-{\hat{I}})=\\
\begin{pmatrix}
\begin{array}{rrrrr}
\hat{V} & -0.48^{+0.20}_{-0.18} & -0.34^{+0.21}_{-0.22} & 0.88^{+0.10}_{-0.25} & 0.60^{+0.11}_{-0.19} \\
\ldots & {\hat{U}}-{\hat{V}} & 0.60^{+0.10}_{-0.13} & -0.38^{+0.18}_{-0.17} & -0.33^{+0.15}_{-0.14} \\
\ldots  & \ldots & {\hat{B}}-{\hat{V}} & -0.21^{+0.17}_{-0.17} & 0.08^{+0.14}_{-0.15} \\
\ldots  & \ldots  &\ldots  & {\hat{V}}-{\hat{R}} & 0.65^{+0.06}_{-0.07} \\
\ldots  & \ldots  &\ldots  &\ldots & {\hat{V}}-{\hat{I}} \\
\end{array}
 \end{pmatrix}.
 \label{cor_intrinsic:eqn}
 \end{multline}
To reduce clutter only the upper triangular part of the otherwise symmetric matrix is shown.
These color variances are smaller than those derived by \citet{2003A&A...404..901N} or \citet{2007ApJ...659..122J}, for cases where
a direct comparison can be made.
\citetalias{2017ApJ...842...93M}  find an intrinsic dispersion of the SALT2 $c$ parameter (associated with ${\hat{B}}-{\hat{V}}$) of $\approx 0.07$ mag,
though variation of dust properties accounts for some of that dispersion.
%model11_rubin.py
Having no intrinsic color dispersion is disfavored, and not allowing for them can have important consequences.  Running an analysis assuming no intrinsic dispersion (replacing the probabilistic intrinsic magnitudes in Eq.~\ref{ewsiv:eqn}
with a strict equality), we find significant shifts in all slope parameters  and 
artificially
smaller statistical uncertainties.

@@

The above result is visualized in Figure~\ref{plane:fig}, which  shows  in the
${\hat{U}}{\hat{V}}\hat{I}$-subspace
two perspectives (left and right panels)
of the unit vectors corresponding to $\vec{\gamma}^0$ and  $\vec{\gamma}^1$ of our model (solid lines),
and $\vec{a}$, $\vec{b}$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  All  vectors are set to intersect the origin.
The two perspectives show that while the four vectors point in different directions for each band combination,
they are almost coplanar in ${\hat{U}}{\hat{V}}{\hat{I}}$ (they are slightly less coplanar in ${\hat{B}}{\hat{V}}{\hat{R}}$).  The $\vec{a}$ and $\vec{b}$ vectors and the $\vec{\gamma}^0$ and $\vec{\gamma}^1$
vectors span near-parallel planes in the 5-dimensional color space, and hence the latter can be almost entirely expressed in terms of
the former.

@@

Given that
$\vec{\gamma_0}$  has nearly the same effect on color as dust extinction with constant $R^F$,
the distribution of
our parameter $\gamma^0_{\hat{V}} k_0$ can be compared with the expected distribution
of $A^F_V$.
Figure~\ref{k0_med:fig} shows the histogram of
per-supernova
median values
from the Monte Carlo chains
and the  stack of the posteriors of all supernovae  for $\gamma^0_{\hat{V}} k_0$ 
relative to that of an arbitrary supernova  $\gamma^0_{\hat{V}} k_0|_0$.
(Subtracting the value of the one supernova suppresses correlated errors in $k_0$ common to all supernovae.)
The distributions are non-Gaussian, having a sharp rise in the blue and an extended tail in the red.  This is consistent
with
simulations based on expected dust distributions
within galaxies and the distribution of galaxy orientations with respect to the observer,
which determine column densities along the lines of sight toward supernovae that are consistent with observations
\citep{1998ApJ...502..177H, 2007ApJ...659..122J}.  
We emphasize that unlike other analyses,
this distribution is given by the data, and so is not dependent on any knowledge or prior
for the distribution of dust extinction the SN population is expected to suffer.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.8in]{deltagamma0_med.pdf}
   \includegraphics[width=2.8in]{deltagamma1_med.pdf}
      \caption{
      Normalized stack of the posteriors of all supernovae  and histogram 
      of  per-supernova 
median values of: (left) $\gamma^0_{\hat{V}} k_0-\gamma^0_{\hat{V}} k_0|_0$, which is associated with the $A^F_V$ of an $R^F \sim 1/\kappa_1$ dust model;
(right) $\gamma^1_{\hat{V}} k_1-\gamma^1_{\hat{V}} k_1|_0$,
which  is associated with extinction corrections due to deviations away from the canonical $\langle R^F_{\mathit{eff}}\rangle$ value.
   \label{k0_med:fig}}
\end{figure}

We now consider the direction of $\vec{\gamma}^1$. 
Different supernovae are expected to encounter dust with different $R^F$;
having  associated $\vec{\gamma}^0$ with
the extinction from a single effective
$\langle R^F_{\mathit{eff}}\rangle$, $\vec{\gamma}^1$ is associated with perturbations
away from that effective value.
Let us introduce a new parameter for each supernova, $\delta R$, such that 
$A^F =\langle R^F_{\mathit{eff}}\rangle(1-  \delta R) E^F(B-V)$.
Variations in the value of  $\delta R$, with the prior of the non-correlation between 
the coefficients in Eq.~\ref{newdust:eqn},
lead 
to 
\begin{equation}
\kappa_2 =\langle R^F_{\mathit{eff}}\rangle \left(1- \frac{\mbox{Var}(\delta R \times E^F(B-V))}{\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))}\right).
\end{equation}
From the matrix elements of Eq.~\ref{trans_I:eqn},  $\kappa_2=-5.27/0.72=-7.31$, so that $\mbox{Var}(\delta R \times E^F(B-V)) = 4.00\times\mbox{Cov}(\delta R \times E^F(B-V), E^F(B-V))$.  This result indicates that there is a small but significant covariance between  extinction contributions due to
fractional deviations from an effective $R^F$ and
the color-excess parameter $E^F(B-V)$.
Figure~\ref{k0_med:fig} shows the histogram of per-supernova
median values
from Monte Carlo chains
and  stack of the posteriors of all supernovae  of $\gamma^1_{\hat{V}} k_1$
relative to that of an arbitrary supernova  $\gamma^1_{\hat{V}} k_1|_0$.

@@

\subsection{Results}

@@

The distributions of both sources of color excess $E({\hat{B}}-{\hat{V}})$ and
 $E_\delta({\hat{B}}-{\hat{V}}) \equiv   (\delta_{\hat{B}}-\delta_{\hat{V}})D$ are shown in Figure~\ref{ebv:fig} as the  stack of the posteriors of all supernovae 
 and the histogram of per-supernova median values.
 (These values are relative
to
 those of an arbitrary supernova in order to null out correlated errors.)
The range of ${\hat{B}}-{\hat{V}}$ colors is $> 5 \times$ larger for the external contribution than for the
intrinsic
component.  
The standard deviations of  the median $E_\gamma({\hat{B}}-{\hat{V}})$ and $E_\delta({\hat{B}}-{\hat{V}})$ are
%-----
0.075
and 0.014
%-----
mag respectively.
It should be noted that
the $E_\delta({\hat{B}}-{\hat{V}})$ standard deviation is comparable to its measurement uncertainty per-supernova.
Thus, since
the color excesses of individual supernovae have low signal-to-noise,
the significance of $\vec{\delta}$
is drawn from the ensemble of objects.
